---
title: "Smokes"
author: "Oluchi Nnadi"
format: html
execute:
  echo: false
---


```{r}
#| message: false
library(tidyverse)
library(primer.tutorials)
library(tidymodels)
library(broom)
```

```{r}
smokes |>
  mutate(age_bin = 5 * (age %/% 5)) |>
  group_by(sex, age_bin) |>
  summarize(share_yes = mean(smoke == "Yes"), .groups = "drop") |>
  ggplot(aes(x = age_bin, y = share_yes, color = sex)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.1) +
  scale_y_continuous(labels = scales::label_percent(),
                     limits = c(0, 1)) +
  facet_wrap(~ sex) +
  labs(
    title    = "Share of Ever-Smokers by Age and Sex",
    subtitle = "Older respondents are more likely to be ever-smokers, especially among men",
    x        = "Age (binned to 5 years)",
    y        = NULL,
    caption  = "Source: 1,000-row NHANES cut via primer.tutorials::smokes"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
fit_smokes <- logistic_reg(engine = "glm") |>
  fit(smoke ~ age + sex, data = smokes)

fit_smokes
```

Summary: Smoking can lead to a plethora of negative long-term health outcomes, and there many factors such as gender and age that make people statistically more likely to smoke on average. Using NHANES data of 1,000 U.S. adults, we will build a model to determine the difference in the probability of being an ever-smoker between a 30-year-old woman and a 70-year-old man. However, it is important to note that society's perspective on smoking has shifted since the collection of the NHANES data, so the effects of gender and age on smoking may be skewed leading to bias in the model.
